---
title: "AN 588 Homework 1"
author: "Akiva Zeff"
date: "2025-01-26"
output: 
  html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***
***
# Module 4 Challenges
***
***
## Challenge 1:
***
First, I create a single string containing the entire sentence.  I decided to go fancy and remove all punctuation:
```{r}
ulysses <- "Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."

# removing punctuation
gsub("[[:punct:]]","",ulysses)
ulysses
```

Next, I create a vector by dividing my string into its individual words using *strsplit( )*.  **However, using *strsplit( )* on its own will create a list of each word.**  I adjust for this by using *unlist( )*, which simplifies the list into a vector:
```{r}
ulysses2 <- unlist(strsplit(ulysses, " ")) # I'm splitting the string at all spaces
ulysses2
```

Finally, I need to extract every third word.  I chose to use the *length( )* function to define the end of the extraction sequence.  This means that it will work as intended even if the length of the original string is increased:
```{r}
ulysses3 <- ulysses2[seq(from = 3, to = length(ulysses2), by = 3)]
ulysses3
```

***
## Challenge 2:
***
I start with the matrix from the module:
```{r}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```

Extracting columns 2, 3, and 6:
```{r}
x <- m[, c(2, 3, 6)] # The comma at the beginning of the bracket allows me to only select for columns
x
```

Extracting rows 6-8:
```{r}
x <- m[c(6:8), ] # Again, the comma specifies that I want rows
x
```

Extracting row 2, column 2 to row 6, column 9:
```{r}
x <- m[c(2:6), c(2:9)]
x
```

***
## Challenge 3:
***
Making the array:
```{r}
a <- array(data = 400:1, dim = c(5, 5, 4, 4))
# "400:1" is descending order
```

Results of various extractions (in separate chunks for ease of running individually):
```{r, eval = FALSE}
a[1, 1, 1, 2]
a
# I chose not to include the results of these chunks within my knit document because they're so long.  Feel free to run them on your own!
```
```{r, eval = FALSE}
a[2, 3, 2, ]
a
```
```{r, eval = FALSE}
a[1:5, 1:5, 3, 3]
a
```

***
## Challenge 4:
***
I work from smallest to largest, start by defining the taxa within Anthropoidea and combining them into a list:
```{r}
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Anthropoidea <- list(Platyrrhini, Catarrhini)
names(Anthropoidea) <- list("Platyrrhini", "Catarrhini")
Anthropoidea
```

I repeat the same procedure for Tarsioidea, and then combine them into a larger list encompassing all of Haplorhini:
```{r}
Tarsioidea <- "Tarsiidae"
Tarsioidea

Haplorhini <- list(Anthropoidea, Tarsioidea)
names(Haplorhini) <- list("Anthropoidea", "Tarsioidea")
Haplorhini
```

For Strepsirhini, I simply use the code given in the module:
```{r}
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")
Strepsirhini
```

Finally, I combine both lists into one large "Primates" list:
```{r}
Primates <- list(Haplorhini, Strepsirhini)
names(Primates) <- list("Haplorhini", "Strepsirhini")
Primates
```

***
## Challenge 5:
***
First, I create the matrix:
```{r}
m3 <- matrix(data = c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow = 5, ncol = 3, byrow = TRUE) # "byrow = TRUE" must be included, as bycol is the default
m3
```

Next, I coerce it into a data frame:
```{r}
df = data.frame(m3)
df
```

I use the *$ operator* to specify that I only want coerce the second column to be logical-valued:
```{r}
df$X2 <- as.logical(df$X2) #"X2" is the default name given to my second column
df
```

I do more or less the same thing to coerce the third column to be factor-valued:
```{r}
df$X3 <- as.factor(df$X3)
df
```

Finally, I check the data type of each variable using *str( )*:
```{r}
str(df)
```
